Reproducible Workflow on a Public Cloud for Computational Fluid Dynamics

### 2019-03-23 tex/MesnardoBarba2019.tex
Revise benchmarks subsections, and others
* Give more details about the OSU Micro-Benchmarks used.
* Add more details about the Poisson tests.

Also:

* Add paragraph on PetIBM.

\begin{table*}[!h]
    \renewcommand{\arraystretch}{1.5}
    \caption{Hardware Specifications of Nodes Used on Microsoft Azure and Colonial One. On Both Platforms, MPI Applications Take Advantage of RDMA (Remote Direct Memory Access) Network with FDR InfiniBand.}
    \label{tab:hw_specs}
    \centering
    \begin{tabular}{cccccccc}
        Platform & Node & Intel Xeon CPU & \# threads & NVIDIA GPU & RAM (GiB) & SSD Storage (GiB) \\
        \hline
        \multirow{2}{*}{Azure} & NC24r & Dual 12-Core E5-2690v3 (2.60GHz) & 24 & 4 x K80 & 224 & 1440 \\
        & H16r & Dual 8-Core E5-2667v3 (3.20GHz) & 16 & - & 112 & 2000 \\
        \hline
        \multirow{2}{*}{Colonial One} & Ivygpu & Dual 6-Core E5-2620v2 (2.10GHz) & 12 & 2 x K20 & 120 & 93 \\
        & Short & Dual 8-Core E5-2650v2 (2.60GHz) & 16 & - & 120 & 93 \\
        \hline
    \end{tabular}
\end{table*}

@@

Thanks to a sponsorship ($\$20,000$ worth of cloud credits) from the Microsoft Azure for Research program,\footnote{Microsoft Azure for Research: \url{https://www.microsoft.com/en-us/research/academic-program/microsoft-azure-for-research/}} we ran three-dimensional CFD simulations of the flow around a gliding-snake model using our in-house research software PetIBM\cite{chuang_et_al_2018}.
Prior to that, we ran benchmarks to confirm we would get comparable performances between a local university-managed HPC cluster (Colonial One) and a public cloud platform such as Microsoft Azure.

@@

We compared the MPI communication performances between Microsoft Azure and Colonial One using the micro-benchmarks from the Ohio State University\footnote{OSU Micro-Benchmarks (version 5.6): \url{http://mvapich.cse.ohio-state.edu/benchmarks/}} to perform point-to-point tests and evaluate the latency and bandwidth on the two clusters.
On Colonial One, we used two nodes (provide details of CPU) with InfiniBand network.
On Microsoft Azure, we used two NC24r instances (provide details of CPU) with RMDA over InfiniBand.
Fig. \ref{fig:osu_benchmarks} report the latencies and bandwidths averaged over 5 repetitions.

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/osu_latency_bandwidth.pdf}
    \caption{Point-to-point latency (top) and bandwidth (bottom) obtained on Colonial One (Ivygpu nodes) and on Microsoft Azure (NC24r nodes).}
    \label{fig:osu_benchmarks}
\end{figure}

@@

We compare the runtimes on the two platforms (Colonial One and Microsoft Azure) to solve a three-dimensional Poisson system (central differences on a uniform mesh) on CPUs and GPUs.
Fig. \ref{fig:poisson_benchmarks} reports the runtimes (averaged over 5 repetitions) to solve the Poisson system.
On the CPU nodes, we perform strong scaling on a mesh size of $1000 \times 1000 \times 50$ cells using up to 8 nodes to solve the system with PETSc using a Conjugate Gradient method and a classical algebraic multigrid preconditioner from the package Hypre BoomerAMG.
On the nodes that feature GPU devices, we perform weak scaling starting with a mesh size of $500 \times 500 \times 25$ cells on a single node (using 12 CPU cores and 2 GPU devices); we double the mesh size as we double the number of nodes in the pool.
We use the NVIDIA AmgX library to solve the system with a Conjugate Gradient method and a classical algebraic multigrid preconditioner.
As the number of iterations to reach convergence increases with the size of the system, we normalize the runtimes by the number of iterations.

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/poisson_time_vs_nodes.pdf}
    \caption{Runtimes to solve a Poisson system on Colonial One and on Microsoft Azure. Benchmark was repeated 5 times and we report the mean runtimes as well as the extrema. Poisson system was solved on Azure using PETSc on H16r nodes and using AmgX on NC24r nodes. On Colonial One, we solved the system with PETSc on Short nodes and with AmgX on Ivygpu nodes. The runs with PETSc solved a Poisson system on a mesh size of 50 million cells (top). With AmgX (bottom left), the base mesh size contains 6.25 million cells and we scale it with the number of nodes. On Azure, we also ran the Poisson benchmark with a finer base mesh size of 25 million cells (bottom right). Runtimes obtained with AmgX were normalized by the number of iterations required to reach a absolute tolerance of $10^{-12}$.}
    \label{fig:poisson_benchmarks}
\end{figure}

@@

>>>>

@@

Thanks to a sponsorship ($\$20,000$ worth of cloud credits) from the Microsoft Azure for Research program,\footnote{Microsoft Azure for Research: \url{https://www.microsoft.com/en-us/research/academic-program/microsoft-azure-for-research/}} three-dimensional CFD simulations with our in-house research software ran on the public cloud platform using the reproducible workflow described in Section \ref{sec:workflow}.
Preliminary results from benchmarks and test-cases showed that we are able to obtain similar performance in terms of latency and bandwidth using the Azure virtual network, comparing to a traditional University-managed HPC cluster (Colonial One).
Table \ref{tab:hw_specs} reports the hardware specifications of the nodes used on Microsoft Azure and Colonial One.

\begin{table*}[!h]
    \renewcommand{\arraystretch}{1.5}
    \caption{Hardware Specifications of Nodes Used on Microsoft Azure and Colonial One. On Both Platforms, MPI Applications Take Advantage of RDMA (Remote Direct Memory Access) Network with FDR InfiniBand.}
    \label{tab:hw_specs}
    \centering
    \begin{tabular}{cccccccc}
        Platform & Node & Intel Xeon CPU & \# threads & NVIDIA GPU & RAM (GiB) & SSD Storage (GiB) \\
        \hline
        \multirow{2}{*}{Azure} & \texttt{NC24r} & Dual 12-Core E5-2690v3 (2.60GHz) & 24 & 4 x K80 & 224 & 1440 \\
        & \texttt{H16r} & Dual 8-Core E5-2667v3 (3.20GHz) & 16 & - & 112 & 2000 \\
        \hline
        \multirow{2}{*}{Colonial One} & \texttt{Ivygpu} & Dual 6-Core E5-2620v2 (2.10GHz) & 12 & 2 x K20 & 120 & 93 \\
        & \texttt{Short} & Dual 8-Core E5-2650v2 (2.60GHz) & 16 & - & 120 & 93 \\
        \hline
    \end{tabular}
\end{table*}

@@

Point-to-point MPI benchmarks from the Ohio State University Micro-Benchmarks\footnote{OSU Micro-Benchmarks (version 5.6): \url{http://mvapich.cse.ohio-state.edu/benchmarks/}} suite ran on Microsoft Azure and Colonial One to investigate performance in terms the latency and bandwidth.
The latency test is carried out in a ping-pong fashion and measures the time elapsed to get a response; the sender sends a message with a certain data size and wait for the receiver to send back the message with the same data size.
The bandwidth test measures the maximum sustained rate that can be achieved on the network; it consists in having the sender sends a fixed number of messages to a receiver that replies only after receiving all of them.
The tests ran on \texttt{NC24r} nodes on Azure and \texttt{Ivygpu} nodes on Colonial One, all them featuring a network interface for RDMA (Remote Direct Memory Access) connectivity to communicate over an InfiniBand network.
Fig. \ref{fig:osu_benchmarks} reports the mean latencies and bandwidths obtained over $5$ repetitions on both platforms.

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/osu_latency_bandwidth.pdf}
    \caption{Point-to-point latency (top) and bandwidth (bottom) obtained on Colonial One (\texttt{Ivygpu} nodes) and on Microsoft Azure (\texttt{NC24r} nodes). Benchmark results are averaged over $5$ repetitions.}
    \label{fig:osu_benchmarks}
\end{figure}

@@

CFD algorithms often require to solve linear systems with iterative methods.
For example, the project method implemented in our research software solve a Poisson system every time step to project the velocity field onto the divergence-free space (to satisfy the incompressibility constraint).
Thus, we investigated the time-to-solution to solve a three-dimensional Poisson system (obtained with a 7-point stencil using central differences) on different nodes of Microsoft Azure and Colonial One.
The system was solved using a Conjugate-Gradient (CG) method with a classical Algebraic MultiGrid (AMG) preconditioner with an exit criterion set to an absolute tolerance of $10^{-12}$.
The iterative solver ran on CPU nodes (\texttt{H16r} instances on Azure and \texttt{Short} nodes on Colonial One) using the CG algorithm from the PETSc library and an AMG preconditioner from Hypre BoomerAMG.
Fig. \ref{fig:poisson_benchmarks} (top) reports the mean runtimes (averaged over $5$ repetitions) to iteratively solve the system, on a uniform grid of $50$ million cells ($1000 \times 1000 \times 50$), as we increase the number of nodes in the pool.
We also solved the Poisson system with the NVIDIA AmgX library on multiple GPU devices using \texttt{NC24r} instances on Azure and \texttt{Ivygpu} nodes on Colonial One.
The Poisson system for a base mesh of $6.25$ million cells ($500 \times 500 \times 25$) is solved on a single node using $12$ MPI processes and $2$ GPU devices; we then double the mesh size as we double the number of nodes.
As the number of iterations to reach convergence increases with the size of the system, we normalize the runtimes by the number of iterations.
Fig. \ref{fig:poisson_benchmarks} (bottom left) shows the normalized mean runtimes ($5$ repetitions) obtained on Azure and Colonial One.
The bottom-right panel of the figure reports the normalized mean runtimes obtained on Azure when we put more constraint on the bandwidth; the base mesh now contains $25$ millions cells ($1000 \times 500 \times 50$).

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/poisson_time_vs_nodes.pdf}
    \caption{Runtimes to solve a Poisson system on Colonial One and on Microsoft Azure. Benchmark was repeated 5 times and we report the mean runtimes as well as the extrema. Poisson system was solved on Azure using PETSc on \texttt{H16r} nodes and using AmgX on \texttt{NC24r} nodes. On Colonial One, we solved the system with PETSc on \texttt{Short} nodes and with AmgX on \texttt{Ivygpu} nodes. The runs with PETSc solved a Poisson system on a mesh size of 50 million cells (top). With AmgX (bottom left), the base mesh size contains 6.25 million cells and we scale it with the number of nodes. On Azure, we also ran the Poisson benchmark with a finer base mesh size of 25 million cells (bottom right). Runtimes obtained with AmgX were normalized by the number of iterations required to reach a absolute tolerance of $10^{-12}$.}
    \label{fig:poisson_benchmarks}
\end{figure}

@@

To study the flow around the flying snake, we develop a in-house CFD software called PetIBM\cite{chuang_et_al_2018}, a open-source toolbox that solves the two- and three-dimensional incompressible Navier-Stokes equations using a projection method (an approximate block-LU decomposition of the fully discretized equations\cite{perot_1993}) and an Immersed-Boundary Method (IBM).
Within this framework, the fluid equations are solved on a extended grid that does not conform to the surface of the immersed body.
To model the presence of the body, the momentum equation is augmented with a forcing term that is activated in the vicinity of the immersed boundary.
This technique allows the use of simple fixed structured Cartesian grids to solve the equations.
PetIBM also implements Immersed Boundary Methods (IBMs) that fit into the projection method and, in the present study, we use an implementation of the IBM proposed in \cite{li_et_al_2016}.
To obtain reproducible computational results, the code used to run the simulations needs to be made available.
In that regard, PetIBM is open source, version-controlled on GitHub,\footnote{PetIBM (version 0.4): \url{https://github.com/barbagroup/PetIBM}} and developed under the permissive BSD-3 clause license.
